---
title: "Chapter 02 -- Exercise13"
author: "Jan Kraemer"
date: "2 May 2019"
output: 
  html_document: default
  github_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(tibble)
library(kableExtra)

knitr::opts_chunk$set(echo = TRUE)
table_colnames <- c("Year", "Fatal accidents", "Passenger deaths", "Death rate")
table_values <-
  tribble(~year, ~fatal_accident_count, ~fatalities_count, ~fatality_rate,
          1976, 24, 734, 0.19,
          1977, 25, 516, 0.12,
          1978, 31, 754, 0.15,
          1979, 31, 877, 0.16,
          1980, 22, 814, 0.14,
          1981, 21, 362, 0.06,
          1982, 26, 764, 0.13,
          1983, 20, 809, 0.13,
          1984, 16, 223, 0.03,
          1985, 22, 1066, 0.15)

```

## The Question


Discrete data: The Table gives the number of fatal accidents and deaths on scheduled
airline flights per year over a ten-year period. We use these data as a numerical example
for fitting discrete data models.

a) Assume that the numbers of fatal accidents in each year are independent with a
    $Poisson(\theta)$ distribution. Set a prior distribution for $\theta$ and determine the posterior
    distribution based on the data from 1976 through 1985. Under this model, give a $95\%$
    predictive interval for the number of fatal accidents in 1986. You can use the normal
    approximation to the gamma and Poisson or compute using simulation.
b) Assume that the numbers of fatal accidents in each year follow independent Poisson
    distributions with a constant rate and an exposure in each year proportional to the
    number of passenger miles flown. Set a prior distribution for $\theta$ and determine the
    posterior distribution based on the data for 1976–1985. (Estimate the number of
    passenger miles flown in each year by dividing the appropriate columns of the Table
    and ignoring round-off errors.) Give a $95\%$ predictive interval for the number of fatal
    accidents in 1986 under the assumption that 8 * 1011 passenger miles are flown that
    year.
c) Repeat (a) above, replacing ‘fatal accidents’ with ‘passenger deaths.’
d) Repeat (b) above, replacing ‘fatal accidents’ with ‘passenger deaths.’
e) In which of the cases (a)–(d) above does the Poisson model seem more or less reasonable? 
    Why? Discuss based on general principles, without specific reference to the
    numbers in the Table.

Incidentally, in 1986, there were 22 fatal accidents, 546 passenger deaths, and a death rate of 0.06 per 100 million miles flown.


### The Table
```{r ex_table}
table_values %>%
  kable(col.names=table_colnames) %>%
  kable_styling(full_width = F)
```


